---
title: "Netflix & Cheese"
---
Crafting the perfect cheeseburger is a lot like creating compelling movies or TV shows here at Netflix. It’s all about balance, creativity, and knowing your audience. Just as a great burger needs the right blend of ingredients, unforgettable show or film demands a strong script, layered characters, engaging visuals, and emotional resonance. You can’t rely on one element alone; everything has to come together to deliver an experience that satisfies and sticks with people. And just like how taste trends evolve, audience expectations shift too, resulting in constantly experimenting, refining, and pushing boundaries to serve up something familiar yet fresh every time.

Today, I will be looking at two different sets of data, the Global top 10 and Countries' Top 10 from the last few years. They will show more insights into the success of our programs here at Netflix. My goal is to find any trends that could allow for future success.

To begin I will look into the Global top 10 data set below.
```{r}
#| echo: false
#| message: false
#| warning: false
if(!require("tidyverse")) install.packages("tidyverse")
library(readr)
library(dplyr)
library(styler)
library(DT)
library(stringr)
if(!dir.exists(file.path("data", "mp01"))){
    dir.create(file.path("data", "mp01"), showWarnings=FALSE, recursive=TRUE)
}

GLOBAL_TOP_10_FILENAME <- file.path("data", "mp01", "global_top10_alltime.csv")

if(!file.exists(GLOBAL_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-global.tsv", 
                  destfile=GLOBAL_TOP_10_FILENAME)
}
GLOBAL_TOP_10 <- read_tsv(GLOBAL_TOP_10_FILENAME)
format_titles <- function(df){
    colnames(df) <- str_replace_all(colnames(df), "_", " ") |> str_to_title()
    df
}
GLOBAL_TOP_10 <- GLOBAL_TOP_10 |>
    mutate(season_title = if_else(season_title == "N/A", NA_character_, season_title))
GLOBAL_TOP_10 |> 
    mutate(`runtime_(minutes)` = round(60 * runtime)) |>
    select(-season_title, 
           -runtime) |>
    format_titles() |>
    datatable(options=list(searching=FALSE, info=FALSE)) |>
    formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
As you could see, each page represents the Global Top 10 for a different week for one of different 4 programs. \
```{r}
#| echo: false
#| message: false
#| warning: false
top_non_eng_film <-
  GLOBAL_TOP_10 |>
  filter(category == "Films (Non-English)") |>
  group_by(show_title) |>
  summarise(weeks_in_top10 = n_distinct(week), .groups = "drop") |>
  arrange(desc(weeks_in_top10)) |>
  slice_head(n = 1)

top_non_eng_film_title <- top_non_eng_film$show_title
top_non_eng_film_weeks <- top_non_eng_film$weeks_in_top10
```
To begin my analysis, I wanted to start with finding out the most succesful Non-English film in the world. It turns out that `r top_non_eng_film_title` was the most successful one, with it being in the top 10 for `r top_non_eng_film_weeks` weeks.

```{r}
#| echo: false
#| message: false
#| warning: false
longest_film <-
  GLOBAL_TOP_10 |>
  filter(str_detect(category, "Films")) |>
  select(show_title, runtime) |>
  distinct() |>
  mutate(runtime_min = suppressWarnings(as.numeric(runtime))) |>
  filter(!is.na(runtime_min)) |>
  arrange(desc(runtime_min)) |>
  slice_head(n = 1)

longest_title <- longest_film$show_title
```
To stay on the films topic, I went to look which was the longest Film to ever appear on the Global top 10. The result was that `r longest_title` was the longest film, with a run time of 224 minutes (Almost 4 hours long).

Next, I want to look into the total global viewership for each of our top program in each category. I computed a data table and got the results below.
```{r}
#| echo: false
#| message: false
#| warning: false
top_by_category <-
  GLOBAL_TOP_10 |>
  group_by(category, show_title) |>
  summarise(total_hours = sum(weekly_hours_viewed, na.rm = TRUE), .groups = "drop") |>
  group_by(category) |>
  slice_max(total_hours, n = 1)

datatable(
  top_by_category,
  caption = "Most Viewed Program per Netflix Global Top 10 Category",
  options = list(pageLength = 5, autoWidth = TRUE)
)
```

It is fair to say that the results are not surprising. All four programs are exceptional, and this is proven by the interaction by customers throughout social media. However, to me, Squid game is the one program that cannot be compared to the rest.\

**Squid Game** is the rare show that turned mass unemployment, existential dread, and childhood nostalgia into a global binge party. What began as a bleak satire on debt and desperation somehow became a worldwide watercooler phenomenon—proof that subtitles don’t scare audiences nearly as much as capitalism does. It has collected over 5 billion views worldwide since its debut. With its candy-colored chaos, moral gut punches, and meme-ready moments, **Squid Game** didn’t just win ratings—it redefined what a global hit could look like. It showed that you don’t need superheroes or Hollywood budgets to dominate Netflix; you just need a green tracksuit, a catchy song, and the collective realization that we’d all probably lose Red Light, Green Light.
```{r}
#| echo: false
#| message: false
#| warning: false
library(lubridate)
df_red_2021 <- GLOBAL_TOP_10 |>
  filter(show_title == "Red Notice") |>
  mutate(week = ymd(week)) |>
  filter(year(week) == 2021)
runtime_min <- 1 * 60 + 58
df_red_2021 <- df_red_2021 |>
  mutate(
    total_view_time_min = weekly_hours_viewed * 60,
    weekly_views = total_view_time_min / runtime_min
  )
views_2021 <- df_red_2021 |>
  summarize(total_views_2021 = sum(weekly_views, na.rm = TRUE)) |>
  pull(total_views_2021)

total_views_2021 <- views_2021 / 1e6
```
The last thing I wanted to do with the Global top 10 data set was to calculate the total number of views a program had in a certain year. For this exercise, I chose the movie Red Notice and finding their views in 2021. After running some code, I calculated that the movie had a total of `r total_views_2021` million views in that year.

Next, I moved on to analyze the country top 10's data set below.
```{r}
#| echo: false
#| message: false
#| warning: false
COUNTRY_TOP_10_FILENAME <- file.path("data", "mp01", "country_top10_alltime.csv")

if(!file.exists(COUNTRY_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-countries.tsv", 
                  destfile=COUNTRY_TOP_10_FILENAME)
}
COUNTRY_TOP_10 <- read_tsv(COUNTRY_TOP_10_FILENAME, na = "N/A")

COUNTRY_TOP_10 |>
  datatable(options=list(searching=FALSE, info=FALSE))
```

In similarity to the other data table, each page represents an individual country's Top 10 for a different week for films or tv shows.
```{r}
#| echo: false
#| message: false
#| warning: false
n_countries <- COUNTRY_TOP_10 |>
  distinct(country_name) |>
  nrow()

total_countries <- n_countries

country_weeks <- COUNTRY_TOP_10 |>
  group_by(country_name) |>
  summarise(total_weeks = n_distinct(week)) |>
  arrange(total_weeks)

outlier_country <- country_weeks |>
  filter(total_weeks < 200)

last_week <- COUNTRY_TOP_10 |>
  filter(country_name %in% outlier_country$country_name) |>
  summarise(last_week = max(week))

ceased_country <- outlier_country$country_name
ceased_timeline <- last_week
```
To begin my analysis, I first want to figure out how many countries around the globe that Netflix operates in since 2021. The result is `r total_countries` countries. However, Netflix does not operate in `r ceased_country` since the week of `r ceased_timeline`.
```{r}
#| echo: false
#| message: false
#| warning: false
tv_data <- COUNTRY_TOP_10 |>
  filter(category == "TV")
tv_data <- tv_data |>
  arrange(country_name, show_title, week)
tv_runs <- tv_data |>
  group_by(country_name, show_title) |>
  mutate(
    week_num = as.integer(as.Date(week)),
    gap = if_else(week_num - lag(week_num, default = first(week_num)) == 7, 0, 1),
    streak_id = cumsum(gap)
  ) |>
  group_by(country_name, show_title, streak_id) |>
  summarise(run_length = n(), .groups = "drop")

longest_run <- tv_runs |>
  arrange(desc(run_length)) |>
  slice(1)

longest_show_run <- longest_run$run_length
longest_country <- longest_run$country_name
longest_show <- longest_run$show_title
```
Switching back to positive things, I wanted to find the TV show who had the longest run inside the top 10 of any country in the world. Turns out, `r longest_show` had a run of `r longest_show_run` consecutive weeks in the top 10 of `r longest_country`. Having watched the show myself, I could understand its popularity given the context as to what happened in real life back in the 80's and 90's.
```{r}
#| echo: false
#| message: false
#| warning: false
debut_weeks <- tv_data |>
  group_by(show_title, season_title) |>
  summarise(debut_week = min(week), .groups = "drop")
debut_countries <- tv_data |>
  inner_join(debut_weeks,
             by = c("show_title", "season_title", "week" = "debut_week"))
country_counts <- debut_countries |>
  group_by(show_title, season_title, week) |>
  summarise(countries_charted = n_distinct(country_name), .groups = "drop") |>
  arrange(desc(countries_charted))
top_debut <- country_counts |>
  slice(1)

debut_show <- paste(top_debut$show_title, top_debut$season_title, sep = " - ")
debut_countries <- top_debut$countries_charted
```
To keep the positivity going, I stayed with TV shows and analyze which one had debuted in the most amount of top 10's around the world. After some analysis, I discovered that `r debut_show` debuted in the top of `r debut_countries` different countries. A very impressive tout. 

Lastly, I analyze TV shows in the US. I began to search which TV shows had not debuted on the US top 10, but eventually made its way onto it. As you can see below, these are the shows that accomplish that feat. The most recent one to do it was KPop Demon Hunters.
```{r}
#| echo: false
#| message: false
#| warning: false
us_films <- COUNTRY_TOP_10 |>
  filter(category == "Films", country_name == "United States")

film_rank_paths <- us_films |>
  group_by(show_title) |>
  summarise(
    debut_week = min(week),
    debut_rank = weekly_rank[which.min(week)],
    max_rank = min(weekly_rank),
    last_week = max(week),
    .groups = "drop"
  )

late_climbers <- film_rank_paths |>
  filter(debut_rank > 1 & max_rank == 1)

late_climbers |>
  arrange(desc(last_week)) |>
  datatable(
    options = list(
      pageLength = 10,
      autoWidth = TRUE
    )
  )
```
###Press Release 1\
```{r}
#| echo: false
#| message: false
#| warning: false
st_data <- tv_data |>
  filter(grepl("Stranger Things", show_title))

st_hours <- sum(st_data$weekly_hours_viewed, na.rm = TRUE)

st_countries <- st_data |>
  distinct(country_name) |>
  nrow()
st_weeks <- st_data |>
  summarise(total_weeks = n_distinct(week)) |>
  pull(total_weeks)
```
### Press Release 1\
**Stranger Things - 5th and Final trip around the sun**  

**Los Gatos, CA — October 2025** — Netflix today announced that the fifth and final season of its global phenomenon *Stranger Things* will premiere at the end of 2025. As anticipation builds, Netflix reflects on the unparalleled success of the series’ first four seasons, which have left an enduring mark on popular culture and television history.  

Across its four seasons, *Stranger Things* has amassed almost 3 billion total amount of hours in viewrship, placing it among the most-watched series ever on the platform. The series appeared in Netflix’s **Global Top 10 for various weeks**, with each season sustaining popularity long after its premiere.  

The show’s reach extends far beyond its hometown of Hawkins, Indiana: *Stranger Things* charted in the Top 10 of `r st_countries`, underscoring its power to connect with audiences across continents. From Brazil to Germany, India to South Korea, the series has proven itself a truly global television event.  

When compared with other blockbuster English-language hits, *Stranger Things* stands in a league of its own. Other shows like *Wednesday*, *The Witcher* or even *Bridgerton* can't matched the sustained, multinational engagement of *Stranger Things*.  

As the curtain prepares to rise on the final season, fans around the globe can look forward to one last adventure with Eleven, Mike, and the beloved Hawkins crew. The conclusion of *Stranger Things* is set to be one of the most significant cultural television moments of the decade, cementing its legacy as a defining series of the streaming era.

### Press Release 2\
**Bollywood goes Streaming**  

**Los Gatos, CA — October 2025** Netflix today highlighted the growing popularity of its Hindi-language films and TV shows in India, the world’s most populated country and one of Netflix’s fastest-growing markets.

Over the past few years, Hindi titles on Netflix have been watched for hundreds of millions of hours, with both films and series regularly landing in the Top 10. This strong demand suggests a large and expanding subscriber base in India, as more viewers turn to Netflix for entertainment in their own language.

Recent hits like RRR show how Netflix is connecting with audiences week after week, with Hindi titles appearing in the Top 10 for various weeks. The steady growth reflects not just short-term excitement but long-term loyalty.

By investing in more Hindi originals and expanding its catalog, Netflix is committed to making India a centerpiece of its global strategy.\

### Press Release 3\
**The rise of Subtitles**  

**Los Gatos, CA — October 2025**— Netflix today announced the rising popularity of non-English language entertainment among U.S. audiences, signaling a major shift in viewing habits and opening new opportunities for international storytelling.

Over the past two years, non-English titles have consistently broken into the U.S. Top 10, from Spanish-language dramas and Korean thrillers to French comedies and Japanese anime. Collectively, these titles have generated hundreds of millions of viewing hours in the U.S., proving that American audiences are increasingly eager to explore stories beyond traditional English-language programming.

Hits such as Squid Game, Money Heist, and Lupin each dominated the U.S. Top 10 for multiple weeks, with Squid Game alone becoming one of Netflix’s most-watched series ever. More recently, non-English films have also surged in popularity, with Hindi, Spanish, and Korean blockbusters drawing strong viewership across the United States.

This trend reflects more than just a cultural curiosity—it highlights a strategic growth opportunity. By expanding investment in dubbing, subtitling, and global marketing, Netflix aims to make international stories even more accessible to American viewers. The company will also continue to partner with creative talent worldwide, ensuring that audiences in the U.S. can experience the diversity of global storytelling.

“Non-English language titles are no longer niche—they are becoming mainstream in the U.S.,” said a Netflix spokesperson. “As we look ahead, we’re excited to bring even more international stories to American homes, strengthening our commitment to entertainment that transcends borders.”\

--------------
Last Updated: `r format(Sys.time(), "%A %m %d, %Y at %H:%M%p")`