---
title: "NYC Tree Party"
format: html
---
New York City’s trees are one of our most valuable public resources — cooling our streets, improving air quality, and strengthening the livability of our neighborhoods. But not every community benefits equally. As a City Council member, I want to ensure that historically overlooked districts receive the same green investment as the rest of the city.

DPR manages nearly 900,000 trees across 30,000 acres of parkland, yet the data shows clear differences in tree density, health, and species diversity across council districts. By analyzing the NYC TreeMap dataset, I aim to highlight where these gaps exist and use that evidence to advocate for targeted funding to expand and maintain our urban forest.

Investing in trees isn’t just beautification — it’s an issue of equity, public health, and resilience. Every New Yorker deserves clean air, shaded streets, and access to the benefits our city’s trees provide.

First, let's take a look at the data and look at all the city council districts by borough.

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(glue)
library(sf)

if(!dir.exists(file.path("data", "mp03"))){
  dir.create(file.path("data", "mp03"), showWarnings = FALSE, recursive = TRUE)
}

library <- function(pkg){
  pkg <- as.character(substitute(pkg))
  options(repos = c(CRAN = "https://cloud.r-project.org"))
  if(!require(pkg, character.only = TRUE, quietly = TRUE)) install.packages(pkg)
  stopifnot(require(pkg, character.only = TRUE, quietly = TRUE))
}

get_nyc_council_districts <- function(simplify_tol_m = NULL){

  zip_url <- "https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/city-council/nycc_25c.zip"

  zip_filename <- basename(zip_url)

  zip_path <- file.path("data", "mp03", zip_filename)
  unz_dir  <- file.path("data", "mp03", tools::file_path_sans_ext(zip_filename))

  if(!file.exists(zip_path)){
    message("Downloading: ", zip_url)
    utils::download.file(zip_url,
                         destfile = zip_path,
                         mode = "wb",
                         quiet = TRUE)
  } else {
    message("Using existing ZIP: ", zip_path)
  }

  if(!dir.exists(unz_dir)){
    message("Unzipping to: ", unz_dir)
    utils::unzip(zip_path, exdir = unz_dir)
  } else {
    message("Using existing unzipped directory: ", unz_dir)
  }

  shp <- list.files(unz_dir, pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)
  if(length(shp) == 0) stop("No .shp found after unzip. Check ZIP contents.")
  shp <- shp[1]

  dat <- sf::st_read(shp, quiet = TRUE)

  dat <- sf::st_transform(dat, crs = "WGS84")

  if(!is.null(simplify_tol_m)){
    tol_deg <- simplify_tol_m / 111320
    dat <- dat |>
      mutate(geometry = sf::st_simplify(geometry,
                                        dTolerance = tol_deg,
                                        preserveTopology = TRUE))
  }
  dat
}
nyc_council_districts <- get_nyc_council_districts()
```
```{r}
#| echo: false
#| message: false
#| warning: false
library(ggplot2)

nyc_ccd_fast <- get_nyc_council_districts(simplify_tol_m = 5)

ggplot(nyc_ccd_fast) +
  geom_sf(color = "black", fill = NA, size = 0.3) +   
  geom_sf_text(aes(label = CounDist), size = 1.5) +     
  theme_minimal() +
  labs(
    title = "NYC City Council Districts",
    x = "Longitude",
    y = "Latitude"
  )
```
As you could see, there are 51 city council districts in New York City. 

Once we have visualized all the districts, next we are going to visualize the trees that within each one.
```{r}
#| echo: false
#| message: false
#| warning: false
library(httr2)
library(dplyr)

nyc_trees <- function(
  url   = "https://data.cityofnewyork.us/resource/hn5i-inap.geojson",
  limit = 5000L
) {
  mp03 <- file.path("data", "mp03")

  offset <- 0L            
  batch  <- 1L            
  done   <- FALSE
  pages  <- list()        

  while (!done) {
    tree_file <- file.path(mp03, paste0("nyc_trees_", offset, ".geojson"))

    if (!file.exists(tree_file)) {
      request(url) |>
        req_url_query(`$limit` = limit, `$offset` = offset) |>
        req_retry(max_tries = 5) |>
        req_perform() |>
        resp_body_raw() |>
        writeBin(con = tree_file)
    }

    tree_page <- sf::st_read(tree_file, quiet = TRUE)
    n_rows    <- nrow(tree_page)

    if (n_rows == 0) {
      # nothing here -> stop
      done <- TRUE
    } else {
      # only touch planteddate if it actually exists
      if ("planteddate" %in% names(tree_page)) {
        tree_page <- tree_page |>
          mutate(planteddate = as.character(planteddate))
      }

      pages[[batch]] <- tree_page

      # if we get fewer rows than limit, we're at the end
      if (n_rows < limit) {
        done <- TRUE
      } else {
        offset <- offset + limit
        batch  <- batch + 1L
      }
    }
  }
  tree_points <- dplyr::bind_rows(pages)
  tree_points
}
tree_points <- nyc_trees()
```
```{r}
#| echo: false
#| message: false
#| warning: false
tree_points <- st_transform(tree_points,    crs = "WGS84")

ggplot() +
  # Draw tree points FIRST (background)
  geom_sf(
    data = tree_points,
    color = "darkgreen",
    alpha = 0.1,
    size  = 0.01
  ) +
  # Draw district boundaries SECOND (on top)
  geom_sf(
    data = nyc_ccd_fast,
    fill  = NA,
    color = "black",
    linewidth = 0.5
  ) +
  coord_sf() +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "NYC Street Trees and City Council Districts",
    x = NULL, y = NULL
  )

```
This plot overlays all NYC street tree locations on top of the City Council district boundaries. Because the city has such a large number of trees, the individual points are so dense that they blend together and form solid green areas rather than distinct dots. This density highlights how heavily planted certain neighborhoods are and gives a quick visual impression of where trees are most concentrated across the city.

Now that we have visualize our council districts and its trees, it's time to do more in depth analysis on the data.

Here will be the 4 things will be looking for

1. Council district with the most trees
2. Council district with the highest density of trees
3. Council district with the highest fraction of dead trees
4. The most common tree species in Manhattan
```{r}
#| echo: false
#| message: false
#| warning: false
library(dplyr)
library(DT)
tree_districts <- st_join(
  tree_points,
  nyc_council_districts,
  join = st_within    
)

trees_by_district <- tree_districts %>%
  st_drop_geometry() %>%      
  group_by(CounDist) %>%
  summarize(n_trees = n()) %>%
  arrange(desc(n_trees))

top10 <- trees_by_district %>% 
  slice_max(n_trees, n = 10) %>%
  rename(`Council District` = CounDist,
         `Tree Count` = n_trees) 
  
datatable(
  top10,
  rownames = TRUE,
  options = list(
    pageLength = 10,
    autoWidth = TRUE
  ),
  caption = "Top 10 NYC Council Districts by Number of Trees"
)
```