---
title: "NYC Tree Party"
format: html
---
New York City’s trees are one of our greatest public assets. They shade our streets, clean our air, and make our neighborhoods more livable — yet not every community benefits equally from this natural infrastructure. As a City Council member, I believe it’s time we address this imbalance by dedicating additional funding to expand and maintain our urban forest, particularly in districts that have been historically overlooked.

The NYC Department of Parks and Recreation (DPR) manages nearly 900,000 trees citywide, covering over 30,000 acres of public parkland. This is a tremendous achievement, but the data shows that some council districts have denser, healthier tree populations than others. To better understand these disparities, I am analyzing the NYC TreeMap dataset to examine how trees are distributed across districts, which species dominate, and how factors like tree health and diversity vary throughout the city.

By grounding this proposal in data, my goal is to build a strong case for targeted investment — a program that not only plants more trees in under-served areas but also ensures long-term care and maintenance. Expanding the city’s tree canopy isn’t just about beautification; it’s about equity, public health, and resilience. Every New Yorker deserves to walk down a shaded block, breathe cleaner air, and share in the environmental and social benefits our trees provide.

First, let's take a look at the data and look at all the city council districts by borough.

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(glue)
library(sf)

if(!dir.exists(file.path("data", "mp03"))){
  dir.create(file.path("data", "mp03"), showWarnings = FALSE, recursive = TRUE)
}

library <- function(pkg){
  pkg <- as.character(substitute(pkg))
  options(repos = c(CRAN = "https://cloud.r-project.org"))
  if(!require(pkg, character.only = TRUE, quietly = TRUE)) install.packages(pkg)
  stopifnot(require(pkg, character.only = TRUE, quietly = TRUE))
}

get_nyc_council_districts <- function(simplify_tol_m = NULL){

  zip_url <- "https://s-media.nyc.gov/agencies/dcp/assets/files/zip/data-tools/bytes/city-council/nycc_25c.zip"

  zip_filename <- basename(zip_url)

  zip_path <- file.path("data", "mp03", zip_filename)
  unz_dir  <- file.path("data", "mp03", tools::file_path_sans_ext(zip_filename))

  if(!file.exists(zip_path)){
    message("Downloading: ", zip_url)
    utils::download.file(zip_url,
                         destfile = zip_path,
                         mode = "wb",
                         quiet = TRUE)
  } else {
    message("Using existing ZIP: ", zip_path)
  }

  if(!dir.exists(unz_dir)){
    message("Unzipping to: ", unz_dir)
    utils::unzip(zip_path, exdir = unz_dir)
  } else {
    message("Using existing unzipped directory: ", unz_dir)
  }

  shp <- list.files(unz_dir, pattern = "\\.shp$", full.names = TRUE, recursive = TRUE)
  if(length(shp) == 0) stop("No .shp found after unzip. Check ZIP contents.")
  shp <- shp[1]

  dat <- sf::st_read(shp, quiet = TRUE)

  dat <- sf::st_transform(dat, crs = "WGS84")

  if(!is.null(simplify_tol_m)){
    tol_deg <- simplify_tol_m / 111320
    dat <- dat |>
      mutate(geometry = sf::st_simplify(geometry,
                                        dTolerance = tol_deg,
                                        preserveTopology = TRUE))
  }
  dat
}
```
```{r}
#| echo: false
#| message: false
#| warning: false
library(ggplot2)

nyc_ccd_fast <- get_nyc_council_districts(simplify_tol_m = 5)

ggplot(nyc_ccd_fast) +
  geom_sf(color = "black", fill = NA, size = 0.3) +   
  geom_sf_text(aes(label = CounDist), size = 3) +     
  theme_minimal() +
  labs(
    title = "NYC City Council Districts",
    subtitle = "District numbers only"
  )
```
As you could see, there are 51 city council districts in New York City. 

Once we have visualized all the districts, next we are going to visualize the trees that within each one.

```{r}
#| echo: false
#| message: false
#| warning: false
library(httr2)
library(dplyr)

nyc_trees <- function(
  url   = "https://data.cityofnewyork.us/resource/hn5i-inap.geojson",
  limit = 5000L
) {
  mp03 <- file.path("data", "mp03")

  offset <- 0L            
  batch  <- 1L            
  done   <- FALSE
  pages  <- list()        

  while (!done) {
    tree_file <- file.path(mp03, paste0("nyc_trees_", offset, ".geojson"))

    if (!file.exists(tree_file)) {
      request(url) |>
        req_url_query(`$limit` = limit, `$offset` = offset) |>
        req_retry(max_tries = 5) |>
        req_perform() |>
        resp_body_raw() |>
        writeBin(con = tree_file)
    }

    tree_page <- sf::st_read(tree_file, quiet = TRUE)
    n_rows    <- nrow(tree_page)

    if (n_rows == 0) {
      # nothing here -> stop
      done <- TRUE
    } else {
      # only touch planteddate if it actually exists
      if ("planteddate" %in% names(tree_page)) {
        tree_page <- tree_page |>
          mutate(planteddate = as.character(planteddate))
      }

      pages[[batch]] <- tree_page

      # if we get fewer rows than limit, we're at the end
      if (n_rows < limit) {
        done <- TRUE
      } else {
        offset <- offset + limit
        batch  <- batch + 1L
      }
    }
  }
  trees_sf <- dplyr::bind_rows(pages)
  trees_sf
}
```
```{r}
#| echo: false
#| message: false
#| warning: false

council_districts <- get_nyc_council_districts(simplify_tol_m = 5)  # or NULL if you want full detail

tree_points <- nyc_trees()

council_districts <- st_transform(council_districts, crs = "WGS84")
tree_points <- st_transform(tree_points,    crs = "WGS84")

ggplot() +
  # council district polygons
  geom_sf(
    data = council_districts,
    fill  = NA,
    color = "grey40",
    linewidth = 0.3
  ) +
  # tree points
  geom_sf(
    data = tree_points,
    color = "darkgreen",
    alpha = 0.2,
    size  = 0.1
  ) +
  coord_sf() +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "NYC Street Trees and City Council Districts",
    subtitle = "2015 Street Tree Census points over City Council district boundaries",
    x = NULL, y = NULL
  )
```
```{r}
#| echo: false
#| message: false
#| warning: false
tree_districts <- st_join(
  tree_points,
  council_districts,
  join = st_within
)
trees_by_district <- tree_districts %>%
  st_drop_geometry() %>%
  group_by(CounDist) %>%
  summarize(n_trees = n()) %>%
  arrange(desc(n_trees))

trees_by_district[1, ]   # district with the most trees
```
```{r}
names(council_districts)
```